A Simple Ratio for Estimating the Effective Reproduction Value of an Epidemic

### 2021-01-16 docs/manuscript/manuscript.Rmd
writing

From a public health perspective, the $R_t$ can be a method to communicate whether an epidemic is growing in a region. However, this communication may be hindered by the fact that many of the current methods for estimating $R_t$ require advanced mathematics to understand and to apply [@cori_new_2013; @wallinga_different_2004; @wallinga_how_2007]. Here, we describe how $R_t$ can be approximated by a ratio of new infections on day $t$ and day $t-d$, where $d$ is the mean generation interval of the infection (Equation 1). The average *generation interval* is defined as the average time from when a primary case becomes infectious to when a secondary case becomes infectious.

@@

As described by previous publications [@wallinga_how_2007], we can define a function $n(a)$ that describes the expected number of new infections created today (time $t$) by a single person who was infected $a$ days ago (at time $t-a$). Therefore, to calculate $I_{ta}$ (the expected number of new infections created today by all individuals who became infected $a$ days ago), we multiply $n(a)$ by the number of new infections created $a$ days ago ($I_{t-a}$).

```{=tex}
\begin{equation}
  I_{ta} = I_{t-a}n(a)
  (\#eq:eq2)
\end{equation}
```

Then to determine the total expected number of new infections today ($I_t$), we sum $I_{ta}$ for all values of $a$. In other words, $I_t$ can be calculated by taking all individuals infected prior to $t$, calculating the expected number of new infections each one will produce based on the age of their infection, and summing all those values together. If we treat $a$ as continuous variables instead of discrete days, we can integrate over all values of $a$, as shown in \cref{eq:renewal-eq-1}.

```{=tex}
\begin{equation} 
  I_t = \int_a^\infty{I_{ta}}da
  (\#eq:renewal-eq-1)
\end{equation}
```
```{=tex}
\begin{equation} 
  I_t = \int_a^\infty{I_{t-a}n(a)}da
  (\#eq:renewal-eq-2)
\end{equation}
```

Next we can substitute the generation interval function ($g(a)$) into \cref{eq:renewal-eq-2} by recognizing that the generation interval distribution is equal to $n(a)$ normalized by the total number of infections a single infected person is expected cause (i.e. $R_t$).

```{=tex}
\begin{equation} 
  g(a) = \frac{n(a)}{R_t}
  (\#eq:generation-interval-eq)
\end{equation}
```
```{=tex}
\begin{equation} 
  I_t = \int_a^\infty{I_{t-a}g(a)R_t}da
  (\#eq:sub-in-rt-1)
\end{equation}
```
```{=tex}
\begin{equation} 
  \frac{I_t}{R_t} = \int_a^\infty{I_{t-a}g(a)}da
  (\#eq:sub-in-rt-2)
\end{equation}
```
```{=tex}
\begin{equation} 
  R_t = \frac{I_t}{\int{I_{t-a}g(a)}da}
  (\#eq:sub-in-rt-3)
\end{equation}
```
If we make the assumption that there is minimal variance in the generation interval distribution (i.e. $g(a)$ equals approximately 1 when $a$ equals the mean generation interval ($d$) and $g(a)$ equals approximately zero for all other values of $a$), then Equation \cref{eq:sub-in-rt-3} can be simplified to:

@@

A deterministic Susceptible-Exposed-Infectious-Resolved (SEIR) model was used to create epidemic simulations. No birth rate, death rate, or R to S transition rates were included. We included `r simulation_parameters$Nsims` simulated epidemics in our analysis. For each epidemic, the mean generation interval of the infection was randomly selected from `r simulation_parameters$mean_gi_range[1]` to `r simulation_parameters$mean_gi_range[2]` days. The generation interval was randomly divided into latent and infectious periods. Each simulation was divided in to `r as.english(simulation_parameters$num_periods[1])` `r simulation_parameters$period_lengths[1]` day periods with differing $R_0$ values (representing changes in social behaviors in the populations); and the $R_0$ in these periods was randomly selected from `r simulation_parameters$r0_range[1]` to `r simulation_parameters$r0_range[2]`. Between each `r simulation_parameters$period_lengths[1]` day period, a transition period was included in which the $R_0$ would gradually change following a logistic curve. The length of the transition period was randomly selected from `r simulation_parameters$transition_lengths[1]` to `r simulation_parameters$transition_lengths[2]` days. Only simulations that lasted at least `r simulation_parameters$min_simLength` and never had less than `r simulation_parameters$min_newCases` cases per day were included in final analysis.

>>>>

From a public health perspective, the $R_t$ can be a method to communicate whether an epidemic is growing in a region. However, this communication may be hindered by the fact that many of the current methods for estimating $R_t$ require advanced mathematics to understand and to apply [@cori_new_2013; @wallinga_different_2004; @wallinga_how_2007]. Typically these methods require a description of the generation interval distribution, which is the distribution of times between when a primary case becomes infectious to when a secondary case becomes infectious.

Here, we describe how $R_t$ can be approximated by a simple ratio of new infections on day $t$ and day $t-d$, where $d$ is the mean generation interval of the infection (\cref{eq:simple-ratio-eq}). Our aim is two-fold: First, by describing a simplified method of calculating $R_t$, it may allow others to easily and rapidly calculate $R_t$. Second, to assess the accuracy of the simple ratio compared to the more advanced methods for calculating $R_t$. This second aim is important, because this simplified ratio has been used to calculate $R_t$ in public health reports but we have not found significant literature describing the performance of the method (at least in English publications).

@@

As described by previous publications [@wallinga_how_2007], we can define a function $n(a)$ that describes the expected number of new cases produced today (time $t$) by a single person who was infected $a$ days ago (at time $t-a$). Therefore, to calculate $I_{ta}$ (the expected number of new infections created today by all individuals who became infected $a$ days ago), we multiply $n(a)$ by the number of new infections created $a$ days ago ($I_{t-a}$).

```{=tex}
\begin{equation}
  I_{ta} = I_{t-a}n(a)
  (\#eq:eq2)
\end{equation}
```
Then to determine the total expected number of new infections today ($I_t$), we sum $I_{ta}$ for all values of $a$, as shown in \cref{eq:renewal-eq-1,eq:renewal-eq-2}. In other words, $I_t$ can be calculated by taking all individuals infected prior to $t$, calculating the expected number of new infections each one will produce based on the age of their infection, and summing all those values together.

```{=tex}
\begin{equation} 
  I_t = \sum_{a=0}^\infty{I_{ta}}da
  (\#eq:renewal-eq-1)
\end{equation}
```
```{=tex}
\begin{equation} 
  I_t = \sum_{a=0}^\infty{I_{t-a}n(a)}da
  (\#eq:renewal-eq-2)
\end{equation}
```
Next we can substitute the generation interval function ($g(a)$) into \cref{eq:renewal-eq-2} by recognizing that the generation interval distribution is equal to $n(a)$ normalized by the total number of infections a single infected person is expected cause (i.e. $R_t$).

```{=tex}
\begin{equation} 
  g(a) = \frac{n(a)}{R_t}
  (\#eq:generation-interval-eq)
\end{equation}
```
```{=tex}
\begin{equation} 
  I_t = \sum_{a=0}^\infty{I_{t-a}g(a)R_t}da
  (\#eq:sub-in-rt-1)
\end{equation}
```
```{=tex}
\begin{equation} 
  \frac{I_t}{R_t} = \sum_{a=0}^\infty{I_{t-a}g(a)}da
  (\#eq:sub-in-rt-2)
\end{equation}
```
```{=tex}
\begin{equation} 
  R_t = \frac{I_t}{\sum{I_{t-a}g(a)}da}
  (\#eq:sub-in-rt-3)
\end{equation}
```
If we make the assumption that there is minimal variance in the generation interval distribution (i.e. $g(a)$ equals approximately 1 when $a$ equals the mean generation interval ($d$) and $g(a)$ equals approximately zero for all other values of $a$), then Equation \cref{eq:sub-in-rt-3} can be simplified to:

@@

A deterministic Susceptible-Exposed-Infectious-Resolved (SEIR) model was used to create epidemic simulations. No birth rate, death rate, or R to S transition rates were included. We included `r simulation_parameters$Nsims` simulated epidemics in our analysis. For each epidemic, the mean generation interval of the infection was randomly selected from `r simulation_parameters$mean_gi_range[1]` to `r simulation_parameters$mean_gi_range[2]` days. The generation interval was randomly divided into latent and infectious periods. Each simulation was divided in to `r as.english(simulation_parameters$num_periods[1])` `r simulation_parameters$period_lengths[1]` day periods with differing $R_0$ values (representing changes in social behaviors in the populations); and the $R_0$ in these periods was randomly selected from `r simulation_parameters$r0_range[1]` to `r simulation_parameters$r0_range[2]`. Between each `r simulation_parameters$period_lengths[1]` day period, a transition period was included in which the $R_0$ would gradually change following a logistic curve. The length of the transition period was randomly selected from `r simulation_parameters$transition_lengths[1]` to `r simulation_parameters$transition_lengths[2]` days. Only simulations that lasted at least `r simulation_parameters$min_simLength` days and never had less than `r simulation_parameters$min_newCases` cases per day were included in final analysis.
